# Comparisons {.unnumbered}

```{r set}
#| message: false
#| warning: false
library(tidyverse)
ducke <- read_tsv("data/derived_data/ducke_sum.tsv") %>% 
  mutate(scientific = paste(genus, species)) %>% 
  left_join(read_tsv("data/derived_data/cor_tab.tsv")) %>% 
  select(family_cleaned, genus_cleaned, species_cleaned, hand, abundance) %>% 
  filter(!is.na(species_cleaned)) %>% 
   rename_all(~ gsub("_cleaned", "", .)) %>% 
  mutate(scientific = paste(genus, species))
paracou <- read_tsv("data/derived_data/paracou_sum.tsv") %>% 
  mutate(scientific = paste(genus, species)) %>% 
  left_join(read_tsv("data/derived_data/cor_tab.tsv")) %>% 
  select(family_cleaned, genus_cleaned, species_cleaned, hand, abundance) %>% 
  filter(!is.na(species_cleaned)) %>% 
   rename_all(~ gsub("_cleaned", "", .)) %>% 
  mutate(scientific = paste(genus, species))
data <- list(Ducke = ducke, Paracou = paracou) %>% 
  bind_rows(.id = "site") %>% 
  select(-scientific) %>% 
  gather(variable, value, -site, -family, -genus, -species) %>% 
  mutate(variable = paste0(variable, "_", site)) %>% 
  select(-site) %>% 
  group_by(variable, family, genus, species) %>% 
  summarise_all(mean) %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  rowwise() %>% 
  mutate(hand_mean = mean(c(hand_Paracou, hand_Ducke), na.rm = TRUE)) %>% 
  mutate(abundance_mean = mean(c(abundance_Paracou, abundance_Ducke), na.rm = TRUE)) %>% 
  arrange(desc(abundance_mean))
```

## Shared species

Ducke has a total of 948 species after quick taxonomic correction against 706 Paracou of which only 223 (20%) are shared.

```{r venn}
#| message: false
#| warning: false
#| fig-cap: Unique and shared species across sites
library(ggvenn)
ggvenn(list('Ducke (72-ha)' = ducke$scientific, 
            "Paracou (93.75-ha)" = paracou$scientific))
```

```{r shared_sp_tab}
t <- bind_rows(
  head(na.omit(data)),
  tail(na.omit(data))
)
knitr::kable(t, caption = "Top 10 most and least abundant shared species.")
```

> Beware the first is palm. But the third is the Amazonian hyperdominant *Eschweilera Coariacea.*

## Abundances

Abundances are slightly correlated across sites.

```{r abund}
#| message: false
#| warning: false
#| fig-cap: Abundances (N/ha) comparisons for the common species between Ducke and Paracou.
data %>% 
  mutate(label = ifelse((genus %in% t$genus && species %in% t$species),
                        paste0(str_sub(genus, 1, 1), ". ", species), "")) %>% 
  ggplot(aes(abundance_Paracou, abundance_Ducke, col = hand_mean)) +
  geom_abline() +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = label)) +
  theme_bw() +
  scale_x_log10() +
  scale_y_log10() +
  scale_color_viridis_c() +
  scale_color_viridis_c("Mean\nHAND\n[m]") +
  ggpubr::stat_cor() +
  xlab("Paracou Mean Abundance [N/ha]") +
  ylab("Ducke Mean Abundance [N/ha]")
```

## Habitats

Mean HAND does not seem correlated across sites, but this probably a too quick and dirty measure of habitat.

```{r hand}
#| message: false
#| warning: false
#| fig-cap: HAND (m) comparisons for the common species between Ducke and Paracou.
m_hand_duc <- median(data$hand_Ducke, na.rm = TRUE)
m_hand_par <- median(data$hand_Paracou, na.rm = TRUE)
data %>% 
  mutate(label = ifelse((genus %in% t$genus && species %in% t$species),
                        paste0(str_sub(genus, 1, 1), ". ", species), "")) %>% 
  ggplot(aes(hand_Paracou, hand_Ducke, col = abundance_mean)) +
  geom_abline() +
  geom_point() +
  ggrepel::geom_text_repel(aes(label = label)) +
  theme_bw() +
  scale_color_viridis_c("Mean\nAbundance\n[N/ha]", trans = "log") +
  ggpubr::stat_cor() +
  xlab("Paracou Mean HAND [HAND, m]") +
  ylab("Ducke Mean HAND [m]") +
  geom_vline(xintercept = m_hand_par, linetype = "dashed") + 
  geom_hline(yintercept = m_hand_duc, linetype = "dashed")
```

57 species are common between Paracou and Ducke with low HAND values and 57 with high HAND values.

```{r venn_hand}
#| fig-cap: Shared species across sites and high or low values of HAND (above or below the site median).
ggvenn(list("Ducke HAND-" = filter(ducke, hand < m_hand_duc)$scientific, 
            "Paracou HAND-" = filter(paracou, hand < m_hand_par)$scientific, 
            "Ducke HAND+" = filter(ducke, hand >= m_hand_duc)$scientific, 
            "Paracou HAND+" = filter(paracou, hand >= m_hand_par)$scientific))
```

```{r high_hand_tab}
data %>% 
  na.omit() %>% 
  filter(hand_Paracou > m_hand_par, hand_Ducke > m_hand_duc) %>% 
  head() %>% 
  knitr::kable(caption = "Common species between Paracou and Ducke with high HAND values.")
```

```{r low_hand_tab}
data %>% 
  na.omit() %>% 
  filter(hand_Paracou < m_hand_par, hand_Ducke < m_hand_duc) %>% 
  head() %>% 
  knitr::kable(caption = "Common species between Paracou and Ducke with low HAND values.")
```
